Fig2

### 2020-02-20 analysis/figure-2.Rmd
save files
supple table 1

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
source("../R/themes.R")

@@

```{r fig2, fig.width=7}

@@

allDEG %>%
  group_by(sex, tissue, comparison) %>%
  summarize(n = n())

p1 <- allDEG %>%
  filter(comparison == "control_bldg") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.text.y = element_blank())  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = " ", y = "Total Differentially Expressed Genes") +
  scale_fill_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  scale_color_manual(values = sexcolors)

p2 <- allDEG %>%
  filter(comparison != "control_bldg") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Sequential Parental Care Stages  ", y = NULL) +
  scale_fill_manual(values = colorscharnew,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")

@@

fig2 <- plot_grid(barplots, mylegend, nrow = 2, rel_heights = c(1,0.1))
fig2
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(kableExtra)
source("../R/themes.R")

@@

```{r allDEGs}

@@

overlappingDEGs <- allDEG %>%
  filter(tissue == "hypothalamus",
         comparison != "control_bldg") %>%
  mutate(group = paste(sex, tissue,comparison, sep = ".")) %>%
  group_by(gene) %>%
  summarize(n = n()) %>%
  arrange(desc(n))
overlappingDEGs

HYPcandidategenes <- c("OXT", "AVP", "GNRH1",  "AR", "POMC", "AGRP",
                       "CRH", "AVPR1A", "AVPR1B", "AVPR2",
                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR")

suppletable1 <- allDEG %>%
  filter(gene %in% HYPcandidategenes) %>%
  group_by(sex, tissue, comparison) %>%
  summarize(genes = str_c(gene, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = genes ) %>%
  select(sex, tissue, control_bldg, lay_inc.d3, inc.d3_inc.d9,
         inc.d9_inc.d17, hatch_n5)  %>%
  arrange(sex, tissue)
suppletable1

write.csv(suppletable1, "../results/suppltable-1.csv", row.names = F)
```

## variance stabilized gene expression  (vsd)

```{r vsd}
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
```

```{r fig2, fig.width=7, fig.height=7}

p1 <- allDEG %>%
  filter(comparison == "control_bldg") %>%
  filter(tissue == "hypothalamus") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         #axis.text.x = element_blank(),
        legend.position = "none",
        strip.text.y = element_blank())  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = " ", y = "Hypothalamic DEGs") +
  scale_fill_manual(values = colorscharnew,
                       name = " ",
                       drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")  +
  ylim(0,4200)

p2 <- allDEG %>%
  filter(comparison != "control_bldg") %>%
    filter(tissue == "hypothalamus") %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_wrap(~sex) +
  theme_B3() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         #axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Sequential Parental Care Stages  ", y = NULL) +
  scale_fill_manual(values = colorscharnew,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")  +
  ylim(0,1150)

@@

plottopgenes <- function(whichgenes, whichtissue, whichsex, mysubtitle, whichtstages){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()

  candidates$treatment <- factor(candidates$treatment, levels = alllevels)

  p <- candidates %>%
    filter(treatment %in% whichtstages) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot() + 
    facet_wrap(~gene, scales = "free_y") + 
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_text(face = "italic")) +
    labs(y = mysubtitle, x = "Parental stage") 
  return(p)

}

hypDEGs <- c("OXT", "AVP", "GNRH1",  "AR")
topgenes <- plottopgenes(hypDEGs, "hypothalamus", c("female", "male"), "Hypothalamic expression", charlevels)

plot_grid(barplots, topgenes, nrow = 2, rel_heights = c(0.4,0.6))

```

```{r suppletable1}
kable(suppletable1)
```

```{r writefiles}
#write_csv(allDEG, "../results/DEGs-char.csv")
```
